---
title: "Preprcessing Data"
output: html_notebook
---
# Prerequisites

Import required libraries
```{r}
library("tidyverse")
library("dplyr")
```

# Data Wrangling

## Read dataset_1

```{r}
# read data and drop ID col
dataset_1 <- read.csv("../raw_data/dataset_1/winemag-data-130k-v2.csv")[-1]
```

## Read dataset_2

Combine all the csv files for dataset_2
```{r}
# function reads csv values
readfile <- function(filename){
  cat("Reading file: ", filename, "...\n", sep = '' )
  return(read.csv(file = filename, header = TRUE))
}

# function merges two dataframes together
merge_dataframes <- function(dataframe_1, dataframe_2){
  return(rbind(dataframe_1, dataframe_2))
}

# funciton merges all csv files in folder to one dataframe
combine_data = function(mypath){
  filenames=list.files(path=mypath, full.names=TRUE)
  datalist = lapply(filenames, function(x) readfile(x))
  return(Reduce(f = function(x,y) merge_dataframes(x,y), x = datalist, accumulate = FALSE))
}

# merge csv data given dataset folder
dataset_2 <- combine_data("../raw_data/dataset_2")
```

## Join Datasets

Join datasets by `title`,  `description`, `price`, `country`, `points` = `rating`. We choose to join by these characteristics, because it would allow us to safely assume that the wine reviews being merged are the same. However, we intentionally left choose not to join based on certain chacteristics like `province` = `region`, because there were some input errors in the data. For example, in `dataset_1` the `province` was "Sicily & Sardinia" where as in `dataset_2` the  `region` was "Sicily &amp; Sardinia".
```{r}
join_wines <- inner_join(dataset_1, dataset_2, by = c("title", "description", "price", "country", "points" = "rating"))
```

Visualize joined data
```{r}
head(join_wines)
```

Clean joined data to remove redundancies
```{r}
join_wines_cleaned <- join_wines %>%
  select(country, description, designation=designation.x, points, price, province, region=region_1, subregion=region_2, taster_name, taster_twitter_handle, title, variety, winery=winery.x, alcohol, category, url, vintage)

head(join_wines_cleaned)
```


Extract relevant information for our use
```{r}
wines <- join_wines_cleaned %>%
  select(title, alcohol, category, vintage, designation, country, province, region, subregion,  variety, winery, price, points, taster_name, taster_twitter_handle)

head(wines)
```


# Write to CSV

### Joined uncleaned Data

Output uncleaned joined data for others to use
```{r}
write_csv(join_wines, "joined_datasets.csv")
```

### Joined Cleaned Data

Outout cleaned joined data for others to use
```{r}
write_csv(join_wines_cleaned, "joined_datasets_cleaned.csv")
```

### Joined Cleaned Data with relevant categories

Output for easy import later
```{r}
write_csv(wines, "wines.csv")
```
