---
title: "Final Report for Intro to Data Science"
output: html_notebook
---
This is where the final project report write-up goes. 

Before you submit, make sure everything runs as expected.

You can add sections as you see fit. Make sure you have a section called "Introduction" at the beginning and a section called "Conclusion" at the end. The rest is up to you!



##Introduction
- Load the `tidyverse, ggplot, and rtweet` packages
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(rtweet)
library(readr)
```


This data set was scraped from WineEnthusiast, a website that reviews and rates many differet types of wines.  

- This dataset includes information of of 130,000 wine reviews with 10 different data fields. 
  

```{r}
wines <- read.csv(file = '../data/winemag-data-130k-v2.csv')[,-1]
```

```{r}
set.seed(19630217)
wine_sample<- sample_n(wines, 1000)
```

Select the provinces based on points  and Select the best province for wine based on the average points of the sample size. 

#find the average number of points across the 1,000 samples
```{r}
wine_per_province <- wine_sample %>% 
  select(province, points) %>% 
  summarise(points = mean(points))
wine_per_province
```


#Find the best province for wine using the average points across the 1,000 samples
```{r}
best_province <- wine_sample %>% 
  group_by(province, points) %>% 
  filter(points > 88.669)
best_province  
```


Rating distribution

```{r}

```






##Conclusion
